---
title: "p8105_hw2_kz2411"
author: "Keming Zhang"
date: "10/7/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(lubridate)
```

## Problem 1
```{r read and clean sheet}
# read and clean Mr. Trash Wheel
wheel_data <- read_excel('./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx',
                         sheet = "Mr. Trash Wheel", range = "A2:N535") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls,digits = 0)
  )
```

```{r}
# read and clean precipitation data for 2018
precipitation_2018_data <- read_excel('./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx',
                                      sheet = "2018 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    year = 2018
  )
```

```{r}
# read and clean precipitation data for 2019
precipitation_2019_data <- read_excel('./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx',
                                      sheet = "2019 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    year = 2019
  )
```

```{r}
# combine dataset and convert month
precipitation_combine_data <- bind_rows(precipitation_2018_data, precipitation_2019_data) %>%
  mutate(
    month = month.name[month]
  )
```

In Mr. Trash wheel dataset, there are `r nrow(wheel_data)` observations and `r ncol(wheel_data)` variables, such as "weight_tons" and "plastic bottles" which are `r class(pull(wheel_data,"weight_tons"))` and `r class(pull(wheel_data,"plastic_bottles"))` respectively.

In the combination dataset, there are there are `r nrow(precipitation_combine_data)` observations and `r ncol(precipitation_combine_data)` variables, which are year, month and total. And their types are `r class(pull(precipitation_combine_data,"year"))`, `r class(pull(precipitation_combine_data,"month"))` and `r class(pull(precipitation_combine_data,"total"))`.

The total precipitation in 2018 is `r sum(pull(precipitation_2018_data,"total"))`.

The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(wheel_data, year == 2019),"sports_balls"))`.


## Problem 2
```{r}
# clean pols-month.csv
pols_month_data <- read_csv(file = './data/fivethirtyeight_datasets/pols-month.csv') %>%
  janitor::clean_names() %>%
  separate(mon, c("year","month","day"), sep = "-") %>%
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year),
    president = prez_gop,
    president = replace(president, president == 1, "gop"),
    president = replace(president, president == 0, "dem"),
  ) %>%
  select(-day)
```

```{r}
# clean the snp.csv
snp_data <- read_csv(file = './data/fivethirtyeight_datasets/snp.csv') %>%
  janitor::clean_names() %>%
  mutate(
    date = as.character(mdy(date))
  ) %>%
  separate(date, c("year","month","day"), sep = "-") %>%
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year),
    year = ifelse(year > 2021, year - 100, year)
  ) %>%
  select(year,month,everything(),-day)
```

```{r}
# clean the unemployment data
# To ensure that key variables take the same values, convert abb to full name
unemployment_data <- read_csv(file = './data/fivethirtyeight_datasets/unemployment.csv') %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_precentage"
  ) %>%
  mutate(
    month = month.name[match(month,month.abb)]
  ) %>%
  janitor::clean_names()
```

```{r}
# join the datasets
combine_data <- left_join(pols_month_data, snp_data, by = c('year','month')) %>%
  left_join(unemployment_data,by = c('year','month'))
```

The dimension of data in pols-month.csv is (`r dim(pols_month_data)`) and the range of years is `r max(pull(pols_month_data, "year")) -  min(pull(pols_month_data, "year"))`. The key variables are year, month and president.

The dimension of data in snap_csv is (`r dim(snp_data)`) and the range of years is `r max(pull(snp_data, "year")) -  min(pull(snp_data, "year"))`. The key variables are year, month and close.

The dimension of data in unemployment.csv is (`r dim(unemployment_data)`) and the range of years is `r max(pull(unemployment_data, "year")) -  min(pull(unemployment_data, "year"))`. The key variables are year, month and unemployment_precentage.

The dimension of data in the combination data is (`r dim(combine_data)`) and the range of years is `r max(pull(combine_data, "year")) -  min(pull(combine_data, "year"))`. The key variables are year, month,  president, close and unemployment_precentage.

## Problem 3
```{r}
#import and clean data
name_data <- read_csv(file = './data/name.csv') %>%
  janitor::clean_names() %>%
  mutate(
    gender = str_to_title(gender),
    ethnicity = str_to_title(ethnicity),
    childs_first_name = str_to_title(childs_first_name)
  ) %>%
  distinct()
```

```{r}
#produce a table for Olivia
olivia_data = filter(name_data,childs_first_name == "Olivia" & gender == "Female") %>%
  pivot_wider(
    year_of_birth,
    names_from = "ethnicity",
    values_from = "rank"
  ) %>%
  janitor::clean_names()
```

```{r}
#find the most popular name among male children
male_name_number <- name_data %>%
  filter(gender == "Male") %>%
  mutate(
    count = as.numeric(count)
  ) %>%
  group_by(childs_first_name) %>%
  summarize(
    count_total = sum(count)
  ) %>%
  arrange(desc(count_total))
head(male_name_number, 5)
```

Choose Ethan as themost popular name among male children over time

```{r}
#produce a table for Ethan
ethan_data = filter(name_data,childs_first_name == "Ethan" & gender == "Male") %>%
  pivot_wider(
    year_of_birth,
    names_from = "ethnicity",
    values_from = "rank"
  ) %>%
  janitor::clean_names()
```

```{r}
#produce a scatter plot
plot_data <- name_data %>%
  filter(year_of_birth == "2016" & ethnicity == "White Non Hispanic" & gender == "Male")

ggplot(plot_data, aes(x = rank, y = count)) +
  geom_point() + 
  labs(
    x = "rank in popularity",
    y = "number of children"
  )
```


