---
title: "p8105_hw2_kz2411"
author: "Keming Zhang"
date: "10/9/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 6,
                      fig.asp = .6,
                      out.width = "90%"
)
library(tidyverse)
library(readxl)
library(lubridate)
```

## Problem 1
```{r read and clean sheet}
# read and clean Mr. Trash Wheel
wheel_data <- read_excel('./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx',
                         sheet = "Mr. Trash Wheel", range = "A2:N534") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls, digits = 0)
  )
wheel_data
```

```{r read an clean data for 2018}
# read and clean precipitation data for 2018
precipitation_2018_data <- read_excel('./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx',
                                      sheet = "2018 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    year = 2018
  )
precipitation_2018_data
```

```{r read and clean data for 2019}
# read and clean precipitation data for 2019
precipitation_2019_data <- read_excel('./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx',
                                      sheet = "2019 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    year = 2019
  )
precipitation_2019_data
```

```{r combine dataset and convert month}
# combine dataset and convert month
precipitation_combine_data <- bind_rows(precipitation_2018_data, precipitation_2019_data) %>%
  mutate(
    month = month.name[month]
  )
precipitation_combine_data
```

In Mr. Trash wheel dataset, there are `r nrow(wheel_data)` observations and `r ncol(wheel_data)` variables, such as "weight_tons" and "plastic bottles" which are `r class(pull(wheel_data,"weight_tons"))` and `r class(pull(wheel_data,"plastic_bottles"))` respectively.

In the combination dataset, there are there are `r nrow(precipitation_combine_data)` observations and `r ncol(precipitation_combine_data)` variables, which are year, month and total. And their types are `r class(pull(precipitation_combine_data,"year"))`, `r class(pull(precipitation_combine_data,"month"))` and `r class(pull(precipitation_combine_data,"total"))`.

The total precipitation in 2018 is `r sum(pull(precipitation_2018_data,"total"))`.

The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(wheel_data, year == 2019),"sports_balls"))`.


## Problem 2
```{r clean pols-month.csv}
# clean pols-month.csv
pols_month_data <- read_csv(file = './data/fivethirtyeight_datasets/pols-month.csv',
                            show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  separate(mon, c("year","month","day"), sep = "-") %>%
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year), #solve the problem when datasets join
    president = prez_gop,
    president = replace(president, president > 0, "gop"), #solve the problem that there exists 2 in prez_gop
    president = replace(president, president == 0, "dem"),
  ) %>%
  select(-day,-prez_dem,-prez_gop) #remove appointed columns
pols_month_data
```

```{r clean the snp.csv}
# clean the snp.csv
snp_data <- read_csv(file = './data/fivethirtyeight_datasets/snp.csv',
                     show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  mutate(
    date = as.character(mdy(date)) #convert Date to character for separation
  ) %>%
  separate(date, c("year","month","day"), sep = "-") %>%
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year),
    year = ifelse(year > 2021, year - 100, year) #fix the problem of function mdy()
  ) %>%
  select(year,month,everything(),-day) #arrange and remove appointed columns
snp_data
```

```{r clean the unemployment.csv}
# clean the unemployment data
# To ensure that key variables take the same values, convert abb to full name
unemployment_data <- read_csv(file = './data/fivethirtyeight_datasets/unemployment.csv',
                              show_col_types = FALSE) %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_precentage"
  ) %>%
  mutate(
    month = month.name[match(month,month.abb)] #convert month abbreviation to full name for datasets join
  ) %>%
  janitor::clean_names()
unemployment_data
```

```{r join the datasets}
# join the datasets
combine_data <- left_join(pols_month_data, snp_data, by = c('year','month')) %>%
  left_join(unemployment_data,by = c('year','month'))
combine_data
```

The data in pols-month.csv is about number of national politicans who are democratic or republican at any given time. The dimension of it is (`r dim(pols_month_data)`) and the range of years is `r max(pull(pols_month_data, "year")) -  min(pull(pols_month_data, "year"))`. The key variables are year, month and president, and their types are `r class(pull(pols_month_data,"year"))`, `r class(pull(pols_month_data,"month"))` and `r class(pull(pols_month_data,"president"))`.

The data in snap.csv is about S&P stock market index. The dimension of it is (`r dim(snp_data)`) and the range of years is `r max(pull(snp_data, "year")) -  min(pull(snp_data, "year"))`. The key variables are year, month and close, and their types are `r class(pull(snp_data,"year"))`, `r class(pull(snp_data,"month"))` and  `r class(pull(snp_data,"close"))`.

The data in unemployment.csv is about precentage of unemployment at any given time. The dimension of it is (`r dim(unemployment_data)`) and the range of years is `r max(pull(unemployment_data, "year")) -  min(pull(unemployment_data, "year"))`. The key variables are year, month and unemployment_precentage, and their types are `r class(pull(unemployment_data,"year"))`, `r class(pull(unemployment_data,"month"))` and `r class(pull(unemployment_data,"unemployment_precentage"))`.

The dimension of data in the combination data is (`r dim(combine_data)`) and the range of years is `r max(pull(combine_data, "year")) -  min(pull(combine_data, "year"))`. The key variables are year, month,  president, close and unemployment_precentage, and their types are the same as those in original datasets.

## Problem 3
```{r import and clean data}
#import and clean data
name_data <- read_csv(file = './data/name.csv',
                      show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  mutate(
    #address the names of a categorical predictor and string variables
    #only capitalize the first letter of every word
    gender = str_to_title(gender),
    ethnicity = str_to_title(ethnicity),
    ethnicity = replace(ethnicity, ethnicity == "Asian And Paci",
                        "Asian And Pacific Islander"),
    ethnicity = replace(ethnicity, ethnicity == "Black Non Hisp", 
                        "Black Non Hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "White Non Hisp", 
                        "White Non Hispanic"),
    childs_first_name = str_to_title(childs_first_name)
  ) %>%
  distinct() #remove duplicate rows
name_data
```

```{r table for Olivia}
#produce a table for Olivia
olivia_data <- filter(name_data, childs_first_name == "Olivia"
                      & gender == "Female") %>%
  pivot_wider(
    ethnicity,
    names_from = "year_of_birth",
    values_from = "rank"
  )
olivia_data
```


```{r table for male}
#produce a table for male
#select 1st name among different ethnicities in different year
male_data <- filter(name_data, rank == 1 & gender == "Male") %>%
  pivot_wider(
    ethnicity,
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )
male_data
```

```{r scatter plot}
#produce a scatter plot
#produce data for scatter plot
plot_data <- name_data %>%
  filter(year_of_birth == "2016" & ethnicity == "White Non Hispanic" 
         & gender == "Male")

#draw the plot
ggplot(plot_data, aes(x = rank, y = count)) +
  geom_point() + 
  labs(
    x = "rank in popularity",
    y = "number of children"
  )
```


